The impact of gas bulk rotation on the Lyman $\alpha$ line

### 2014-05-15 paper/ms.tex
Conclusions uodated

We use the full width at half maximum (FWHM) to quantify the line
broadening. We measure this width from the velocity histogram by
finding the values of the velocities at half maximum. We use linear
interpolation between histogram points to get a more precise value for
this width.

@@

We measure the peak maxima poisition to quantify the transition from
double into single peak. We do this as a function of the viewing angle
$\theta$ and the maximum rotational velocity. The results are
summarized in Figure  \ref{fig:maximumsvsvelocity}. There are two
interesting features that deserve attention. First, for a viewing angle
parallel to the rotationa axis ($\mu\sim 1.0$) the maxima of all
models are similar regardless of the rotational velocity. Second, at a
viewing angle perpendicular to the rotation axis ($\mu\sim 0.0$) there
is a good fraction of models that become single peaked, a feature that
appears more frequently for homogeneously distributed sources when all
the other parameters are equal.

@@

In the right panel of Figure \ref{fig:Nscatt} we observe how the
average number of scatterings decreases for larger rotational
velocities in the case of homogenous source distributions. For instance, for
$\tau_{\rm H}=10^5$ the average number of scatterings decreases by
$61\%$ at $V_{\rm max}=300\kms$ in comparison to the static case. For
the central source distribution the average number of scatterings
$\langle N_{\rm   scatt}\rangle$ changes less than $0.5\%$.

@@

We perform a similar quantification of in our speherical setup for the
scatic spheres. We find that for the central model the number of
scatterings is proportional to the optical depth, with $\langle N_{\rm
  scatt}\rangle= (1.50, 1.00, 0.92)\tau_{\rm   H}$ for optical depth
values of $\tau_{\rm H} = (10^{5}, 10^{6}, 10^{7})$ respectively.
For the homogeneous static sphere we find that $\langle N_{\rm
  scatt}\rangle= (0.99, 0.59, 0.51)\tau_{\rm   H}$, thisrepresents
factors of $(0.66, 0.59, 0.55)$ lower than the centrally  emitted photons.

@@

Our main result is that rotation clearly impacts the \ly line
morphology. Double peaked lines can make transitions into single
peaked lines as the rotational velocity increases. In the case of
off-centered emission we obtain triple peaked lines for certain
values of optical depth and rotational velocities.

@@

\item In spite of the asymmetry induced by rotation the spatial
  anisotropy in the emission is usually below the
  $10\%$ level. This justifies our consideration of angle-averaged
  effects to quantify the emission line of rotating clouds.

\item The line width increases with rotational velocity. The line width
  approximately follows the functional for  ${\rm FWHM}^2 = {\rm FWHM}_{
    0}^2 + (V_{\rm max}/\lambda)^2$ where FWHM$_{0}$ indicates the line
  width for the static case and $\lambda$ is a constant determined from
  the radiative transfer results it is $\lambda_{\rm c}=3.89\pm0.04$ and
  $\lambda_{\rm h}=1.98\pm0.06$ for the central and homogeneous source
  distribution, respectively.

\item A single peaked line emerges for high rotational velocities in
  the case of homogeneously distributed sources. These cases occur when
  the rotational velocity is larger than half FWHM$_0$.

@@

Comparing our results with recent observed LAEs we find that many 
morphological features such as high central line flux, single peak
profiles  could be explained by gas bulk  rotation present in these
LAEs.

>>>>

We use the full width at half maximum (FWHM) to quantify the line
broadening. We measure this width from the velocity histogram by
finding the values of the velocities at half maximum. We use lineal
interpolation between histogram points to get a more precise value for
this width.

@@

We measure the peak maxima position to quantify the transition from
double into single peak. We do this as a function of the viewing angle
$\theta$ and the maximum rotational velocity. The results are
summarized in Figure  \ref{fig:maximumsvsvelocity}. There are two
interesting features that deserve attention. First, for a viewing angle
parallel to the rotational axis ($\mu\sim 1.0$) the maxima of all
models are similar regardless of the rotational velocity. Second, at a
viewing angle perpendicular to the rotation axis ($\mu\sim 0.0$) there
is a good fraction of models that become single peaked, a feature that
appears more frequently for homogeneously distributed sources when all
the other parameters are equal.

@@

In the right panel of Figure \ref{fig:Nscatt} we observe how the
average number of scatterings decreases for larger rotational
velocities in the case of homogeneous source distributions. For instance, for
$\tau_{\rm H}=10^5$ the average number of scatterings decreases by
$61\%$ at $V_{\rm max}=300\kms$ in comparison to the static case. For
the central source distribution the average number of scatterings
$\langle N_{\rm   scatt}\rangle$ changes less than $0.5\%$.

@@

We perform a similar quantification of in our speherical setup for the
scatic spheres. We find that for the central model the number of
scatterings is proportional to the optical depth, with $\langle N_{\rm
  scatt}\rangle= (1.50, 1.00, 0.92)\tau_{\rm   H}$ for optical depth
values of $\tau_{\rm H} = (10^{5}, 10^{6}, 10^{7})$ respectively.
For the homogeneous static sphere we find that $\langle N_{\rm
  scatt}\rangle= (0.99, 0.59, 0.51)\tau_{\rm   H}$, this represents
factors of $(0.66, 0.59, 0.55)$ lower than the centrally  emitted photons.

@@

Our main result is that rotation clearly impacts the \ly line
morphology. The morphology of the observed spectra for observers at different
viewing angle $\theta$ changes while the flux remains constant for all $\theta$,
for $\mu \sim 1$ double peaked lines can make transitions into single
peaked lines as the rotational velocity increases.

@@

\item Rotation induce a considerable asymmetry in the morphology 
 of the spectra when the viewing angle $\theta$. This occurs for
 all of our models, in particular models with $\mu\sim 1$ i.e for 
 observes aligned with the axis of rotation the rotation does not 
 produce a significance effect while for $\mu\sim 0$ the effect in 
 the FWHM, and Maxima is significant and the number of scatterings.

%\item In spite of the asymmetry induced by rotation the spatial
%  anisotropy in the emission is usually below the
%  $10\%$ level. This justifies our consideration of angle-averaged
%  effects to quantify the emission line of rotating clouds.

\item For $\mu \sim 0$ The line width increases with rotational velocity. The line width
  approximately follows the functional for  ${\rm FWHM}^2 = {\rm FWHM}_{
    0}^2 + (V_{\rm max}/\lambda)^2$ where FWHM$_{0}$ indicates the line
  width for the static case and $\lambda$ is a constant determined from
  the radiative transfer results it is $\lambda_{\rm c}=XX \pm XX$ and
  $\lambda_{\rm h}=XX\pm XX$ for the central and homogeneous source
  distribution, respectively.

\item A single peaked line emerges for high rotational velocities in
  the case of homogeneously distributed sources. These cases occur when
  the rotational velocity is larger than half FWHM$_0$ and $\mu\sim 0$.

@@

%Comparing our results with recent observed LAEs we find that many 
%morphological features such as high central line flux, single peak
%profiles  could be explained by gas bulk  rotation present in these
%LAEs.
